My title

### 2024-10-31 paper/paper.qmd
1. citations in data part done. 2. modified data part in paper 3. file in /script used for prediction data renamed

library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(kableExtra)
library(bayesplot)
```

@@

Polling data for the 2024 U.S. presidential election is provided by FiveThirtyEight [@fivethirtyeight]. This dataset compiles information from numerous polls conducted by various pollsters, detailing public support levels for presidential candidates. This data is periodically updated on FiveThirtyEight's website to reflect the latest polling information for the 2024 election. Each record includes information we need like state in which the poll was conducted (if applicable), polling dates (start and end), the specific pollster, the sample size of each poll, the polling methodology (such as phone, online, etc.), and support percentages for each candidate. This dataset serves as the foundation for analyzing trends and building predictive models for the winner of the upcoming US presidential election. Table tbl-pollingdata below previews this information for polls conducted in 2023 and 2024.

To simulate, test, analyze, and clean the polling data, the statistical programming language R was employed [@citeR]. Key libraries that supported the data analysis include `tidyverse` [@tidyverse], `readxl` [@readxl], `dplyr` [@dplyr], `knitr` [@knitr], `ggplot2` [@ggplot2], and `broom` [@broom] for tidying model outputs. Additionally, the `rstanarm` [@rstanarm] library was utilized to implement multilevel modeling for prediction purposes. Also this analysis inherits the folder structure from Professor Rohan Alexander from University of Toronto [@rohan].

@@

Next, polling organizations develop survey questions based on the desired data, considering sample size, sampling methods, and survey modes to capture respondents' preferences and voting intentions. Different methodologies impact the results, as they reflect the views of distinct groups, which may be influenced by factors like access channels, demographics, and technology usage. For instance, face-to-face surveys tend to capture opinions from offline-inclined groups, whereas voluntary online surveys exclude those less active online【cite】.

Once the survey is designed, polling begins, and responses are aggregated by the polling organization. Respondents' answers are recorded and categorized, converting their opinions into structured data points, though these lack the nuances of human complexity. To protect privacy, data are often anonymized and aggregated, retaining key insights. After collection, raw data undergo cleaning and processing to minimize sampling biases and errors. Yet, any single poll can have random and systematic error sources, so combining multiple poll results helps mitigate these errors【cite】.

Processed data are entered into structured datasets. Ideally, well-formatted polling data are stored as individual records, typically including fields like polling organization, sample size, survey dates, margin of error, candidate support percentages, and methodology description. On platforms like FiveThirtyEight, these fields are standardized for ease of comparison and analysis. Finally, FiveThirtyEight and similar platforms aggregate these structured records and present them to the public or allow non-commercial research use, with data usually updated as new polls are released. This enables analysts and the public to track changes in candidate support, analyze trends, and interpret results. An interesting phenomenon, however, is that as election day approaches, poll results often "converge" or "follow" the average, reducing polling error, with evidence suggesting some "final adjustments" in methodology may occur【cite】.

@@

In researching the likelihood of Trump winning, we selected specific variables to focus on those most impactful for predicting approval rates. Here is an overview of data used for this analysis @tbl-view. Following are selected ourcome variables and predictor variables, and explainations respectively.

### Outcome variables
In this analysis of Donald Trump’s support rate in the 2024 U.S. presidential election, the outcome variable is Trump’s support rate (pct) because it is directly related to the election results. The primary research goal of this analysis is to predict Trump’s chances of winning in the 2024 election, so the support rate is the main metric we aim to model and forecast. The support rate represents the proportion of respondents in a specific poll who support Trump. We seek to understand how the support rate is influenced by different factors, such as polling methodology, time, or geographical variables. By modeling the fluctuations in Trump’s support rate, we can infer his electoral performance, analyze the trends within his support base, and assess the impact of various factors on his popularity. Understanding these dynamics provides insights into Trump’s standing relative to his opponents and supports our overall election prediction.

### Predictor variables
In this analysis, the predictor variables are the factors that we believe can influence the support rate for Trump. These variables are chosen based on their potential to explain changes in support rate. Following are the predictor variables that will be included in our model, along with an explanation of their significance.

**start_date**: Public sentiment and approval rates fluctuate dynamically during an election period, influenced by the passage of time, major events, and campaign strategies. By recording the date of each poll, we can observe changes in support over time.

@@

**methodology**: Different survey methods (e.g., telephone interviews, online questionnaires) significantly impact the representativeness of the results, as they directly influence the characteristics of the sampled population. Some methods may introduce biases toward specific demographic groups; for example, telephone surveys may skew toward older demographics, while online surveys may skew toward younger ones.

@@

**days_since_Biden_Withdrawal**: Trump's main opponent, Kamala Harris, entered the race only after Biden withdrew. Major political events like Biden’s withdrawal can significantly impact voter approval rates, and tracking the time since this event helps assess both its short-term and long-term effects.

Other potential variables (such as voter demographics like gender, age, race, and income) could indeed influence voter approval rates, but they primarily pertain to a **micro-level** analysis. After incorporating GDP and per capita income data from various U.S. states and nationwide in 2023 into a generalized linear model, we found that their correlation with presidential election outcomes was not very strong, likely due to data limitations in capturing the effect of economic changes on the election. Thus, we did not include these factors. Additionally, individual voter characteristics (such as gender and education level) are not consistently available in most polls. In comparison, the variables we chose are more complete and comprehensively capture the macro nature of the data. Therefore, selecting these variables allows us to maintain data simplicity while maximizing the capture of key factors influencing approval rates, avoiding the noise introduced by excessive micro-level variables.

@@

### Figures

@@

@fig-pollstermetho shows the average support rates for the top and bottom ten pollsters, highlighting each pollster’s most commonly used methodology. We can observe that the support rates between the top 5 and bottom 5 pollsters differ by about 10%, which helps us understand the bias introduced by certain pollsters’ influence on support rates, as some pollsters may tend to show higher or lower support rates. This chart also examines the different methodologies used by these pollsters, as varying survey methods (e.g., phone surveys or online panels) may affect the results by reaching different groups of voters. Focusing on these differences aids in assessing the reliability and bias of the data.

@fig-state compares the average support rates across states with the national average, showing regional differences in voter preferences. Due to the Electoral College system in the U.S. presidential election, variations in state support rates can significantly impact the election outcome. This analysis highlights the differences between state and national polls, helping us understand Trump’s relative support in each state. Paying attention to geographic disparities improves the model's prediction accuracy.

@fig-overtime displays the trend in support rates over time for the top five pollsters with the most observations. We chose these pollsters because a sufficient number of observations is needed to accurately capture trend changes and reduce the impact of extreme values. Each subplot represents a pollster and shows the change in Trump’s support rate since Biden’s withdrawal. This helps us observe the dynamic change in support rates following major events, and we can also see how different pollsters respond to the same event. Analyzing time trends aids in understanding shifts in public opinion, providing more time-sensitive support rate predictions.

```{r, message=FALSE, fig.width=5, fig.height=3}
#| label: fig-pollstermetho
#| fig-cap: Support Rates for Donald Trump of Ten Pollsters with Their Most Common used Metholodogy
#| echo: false

@@

ggplot(selected_pollster_data, aes(x = reorder(paste(pollster, "(", count, ")", sep = ""), -avg_pct), y = avg_pct, fill = methodology)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ category, scales = "free_x") +
  labs(
    title = "Top 5 and Bottom 5 Avg Support Rate for Donald Trump",
    subtitle = "With Most Common Used Methodologies of Each Pollsters",
    x = "Pollster (Occurrence Count)",
    y = "Average Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 5, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 6),
    plot.title = element_text(size = 8, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 7, color = "#333333"),
    axis.title = element_text(size = 7, color = "#333333")
  ) +
  scale_fill_brewer(palette = "Set3") # assign colors to different methodologies

```
```{r, message=FALSE, fig.width=5, fig.height=3}
#| label: fig-state
#| fig-cap: Average Support Rate for Trump of Each Category Based on State
#| echo: false

@@

ggplot(selected_state_data, aes(x = reorder(paste(state, "(", count, ")", sep = ""), -avg_pct), y = avg_pct, fill = state)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ category, scales = "free_x") +
  labs(
    title = "Average Support Rate for Doanld Trump by states",
    subtitle = "National Average Support Rate Included",
    x = "State Names or National (Occurrence Count)",
    y = "Average Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 5, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 6),
    plot.title = element_text(size = 8, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 7, color = "#333333"),
    axis.title = element_text(size = 7, color = "#333333")
  ) +
  scale_fill_brewer(palette = "Set3") # assign colors to different states
```

```{r, message=FALSE, fig.width=5, fig.height=3}
#| label: fig-overtime
#| fig-cap: Trend for Average Support Rate for Donald Trump Over Time Since Joe Biden's Withdrawal on 2024 July 21
#| echo: false

# Identify the top 10 pollsters by occurrence
top_pollsters2 <- analysis_data %>%
  count(pollster, sort = TRUE) %>%
  top_n(5, n) %>%
  pull(pollster)

@@

# Step 2: Create the line plot with background distribution
ggplot(support_trend_5day, aes(x = grouped_days, y = avg_pct)) +
  geom_line(aes(color = pollster), linewidth = 1.2) + # Main line color
  geom_jitter(data = data, aes(x = days_since_Biden_Withdrawal, y = pct), 
              color = "#a6cee3", alpha = 0.4, width = 1, height = 0.5) +
  scale_x_continuous(
    breaks = seq(min(analysis_data$days_since_Biden_Withdrawal), 
                 max(analysis_data$days_since_Biden_Withdrawal), 
                 by = 14)
  ) +
  labs(
    title = "Average Support Rate for Donald Trump Over Time",
    subtitle = "Days counted weekly since Biden's Withdrawal",
    x = "Days Since Biden's Withdrawal",
    y = "Average Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 5, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 8, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 7, color = "#333333"),
    axis.title = element_text(size = 7, color = "#333333")
  ) +
  facet_wrap(~ pollster, scales = "free_x")
```

# Model

@@

### Model justification
#### Expectation of relationships between variables
Our model applies a hierarchical Bayesian approach, based on the framework proposed by Gelman et al. (2020). This method builds a layered structure that combines polling data collected by different organizations using various methodologies with regional and temporal trends to more accurately predict support rates on Election Day. Running within a Bayesian framework, our model employs a random walk prior to reflect the dynamically changing support rates following Biden's withdrawal in 2024. This prior distribution is based on historical election trends and the latest polling data, allowing the model to continuously update as new data becomes available.

We anticipate that different polling `methodology` may affect the estimation of support rates. While we cannot precisely determine which polling method might lean toward which demographic group, it is plausible to assume that methods like "online panels" or "phone surveys" might reach different segments of voters, thus introducing some bias into the support rate estimates. For instance, online surveys may engage more frequent internet users, while phone surveys may cover more traditional media users. This bias could lead to varying estimates of support for Trump. By introducing the methodology variable into the model, we aim to correct for these systematic biases associated with different polling methods. Specifically, by controlling for the estimation results across different methods, the model can isolate the biases introduced by each method, allowing the final support rate estimate to better reflect the preferences of the overall voter population. This way, the model can capture potential differences across methods when incorporating data from various polling sources, thereby reducing the systematic error caused by these differences.

For the variable `days_since_Biden_Withdrawal`, we hypothesize that, over time, Trump’s approval ratings may fluctuate due to shifts in the election landscape, policy discussions, and changes in voter sentiment. This hypothesis is based on the dynamic nature of polling data: as elections near their end, poll results typically converge, reflecting more accurate forecasts for election day (see Blumenthal's Polls, Forecasts, and Aggregators). The Economist's forecast model explains this trend, noting that as election day approaches, voters are more likely to solidify their choices, and polling results tend to capture actual public opinion more accurately while the effects of short-term dynamics diminish (How This Works, The Economist). The time factor not only captures shifts in approval ratings as election day draws closer but also allows us to observe how key political events impact approval ratings. For example, candidate debates, policy announcements, or major international events could significantly influence voter preferences at various points in time (see Blumenthal 2010). Thus, we anticipate that the `days_since_Biden_Withdrawal` variable may exhibit a positive or negative trend, depending on the political climate and voter responses to Trump in particular periods.

Through the `is_national` variable, we distinguish between national and state-level polls. We expect that national polls, which typically have larger and more representative samples, are better suited to capturing overall trends, while state-level polls may reflect finer, localized biases and specific regional issues. Because national polls often encompass a broader and more balanced voter base, they provide a more stable forecast of overall voter trends. In contrast, due to geographic and demographic variations, state-level polls often exhibit greater fluctuations in error than national polls. These fluctuations stem from regional issues and diverse demographic characteristics, which may not align fully with national trends (Disentangling Bias and Variance in Election Polls). By introducing `is_national`, this variable enables the model to capture support rate differences arising from geographic and demographic differences. For example, Trump’s support may be higher in traditionally red states, while in swing states, where competition is intense and voter interest is high, his support rates may fluctuate more significantly. This approach helps explain the regional differences in support rates influenced by political leanings and campaign dynamics.

As noted above, differences in state political backgrounds may lead to systematic shifts in support rates; therefore, we included a random intercept `(1 | state)` for each state, with the intention that this random intercept would capture the differing impacts each state has on Trump’s support rate. Similarly, political biases and methodological differences across polling organizations may affect their support rate results. By incorporating `(1 + days_since_Biden_Withdrawal | pollster)`, we allow for variability in estimates by different pollsters over time. We anticipate that, due to differences in the stance or methodologies of these organizations, their estimates of Trump’s support rate in the period following Biden’s withdrawal will vary accordingly.

# Results
## Model Summary

Our results are summarized in a table which is available [here](https://8360846bf92342d59862bf4ec3515643.app.posit.cloud/file_show?path=%2Fcloud%2Fproject%2Fpaper%2Fmodel_summary_table.html). This summary reveals the effects of different polling methods and organizations on our estimates. For instance, there is a small difference in estimates between data collected via online panels and phone surveys. The model adjusts for these differences by using a hierarchical structure, which accounts for inconsistencies caused by methodological and organizational biases. Specifically, variance parameters for different methods (such as online ads vs. phone surveys) indicate the different impacts these methods have on sampling coverage and voter behavior, helping the model balance these differences.

@@

modelsummary::modelsummary(
  list("First model" = first_model),
  statistic = "mad",
  fmt = 2,
  output = here::here("/cloud/project/paper/model_summary_table.html")  # Save as HTML file
)
```

## Prediction Output
From @fig-results-1, the model predicts that nationwide support for Trump on Election Day, November 5, 2024, will center between 45% to 50% and shows moderate fluctuations. This variability is mainly attributed to national events and differences in polling methods. For instance, nationwide events such as economic shifts or unexpected political occurrences may influence voter sentiment, while differences in polling methods (such as phone surveys versus online panels) may introduce slight fluctuations in estimates due to differences in sample coverage.

At the state level, @fig-results-2 shows predictions indicate significant variations between states. For example, support in Texas is considerably higher than in California, with a gap of over 10 percentage points, while swing states like Pennsylvania and Michigan display greater volatility. This result further validates the model’s robustness; the introduction of a random intercept for each state reflects systematic differences across states. This design aligns with data trends, suggesting that states with diverse geographic and demographic profiles exhibit unique political tendencies.

```{r, fig.width=5, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-results
#| fig-cap: "TODO"
#| warning: false
national_only <- prediction_data[prediction_data$state == "national", ]
all_states <- prediction_data[prediction_data$state != "national", ]

ggplot(national_only, aes(x = state, y = predicted_support)) +
  geom_point(size = 3, color = "lightblue") +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, color = "lightblue") +
  labs(title = "Election Day National Trump Support Rate Prediction", y = "Predictied Support Rate (%)", x = "Area Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 6),
    plot.title = element_text(size = 8, face = "bold", color = "#333333"),
    axis.title = element_text(size = 7, color = "#333333"))

# Plot for state data
ggplot(all_states, aes(x = state, y = predicted_support)) +
  geom_jitter(size = 3, color = "lightblue", width = 0.3) +  # Adds slight horizontal jitter to reduce overlap
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, color = "lightblue", alpha = 0.6) +
  geom_smooth(aes(group = 1), method = "lm", color = "orange", se = FALSE) +  # Adds a trend line across all states
  geom_hline(yintercept = 50, color = "black", linetype = "dashed", size = 0.7) +  # Adds a reference line at 50%
  scale_y_continuous(breaks = seq(0, 70, by = 10)) +  # Formats y-axis as percentages with 1 decimal precision
  labs(title = "Election Day Trump Support Rate Prediction by State", y = "Predictied Support Rate (%)", x = "States") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6),
    plot.title = element_text(size = 8, face = "bold", color = "#333333"),
    axis.title = element_text(size = 7, color = "#333333"))

```

>>>>

library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(kableExtra)
library(bayesplot)
library(patchwork)
```

@@

Polling data for the 2024 U.S. presidential election is provided by FiveThirtyEight [@fivethirtyeight]. This dataset compiles information from numerous polls conducted by various pollsters, detailing public support levels for presidential candidates. This data is periodically updated on FiveThirtyEight's website to reflect the latest polling information for the 2024 election. Each record includes information we need like state in which the poll was conducted (or conducted nationally), polling dates (start and end), the specific pollster, the sample size of each poll, the polling methodology, and support percentages for each candidate. This dataset serves as the foundation for analyzing trends and building predictive models for the winner of the upcoming US presidential election. Table tbl-pollingdata below previews this information for polls conducted in 2023 and 2024.

To simulate, test, analyze, and clean the polling data, the statistical programming language R was employed [@citeR]. Key libraries that supported the data analysis include `tidyverse` [@tidyverse], `readxl` [@readxl], `dplyr` [@dplyr], `knitr` [@knitr], `ggplot2` [@ggplot2], `kableExtra` [@kableExtra] and `patchwork` [@patch]. Additionally, the `rstanarm` [@rstanarm] library was utilized to implement multilevel modeling for prediction purposes. Also this analysis inherits the folder structure from Professor Rohan Alexander from University of Toronto [@rohan].

@@

Next, polling organizations develop survey questions based on the desired data, considering sample size, sampling methods, and survey modes to capture respondents' preferences and voting intentions. Different methodologies impact the results, as they reflect the views of distinct groups, which may be influenced by factors like access channels, demographics, and technology usage. For instance, face-to-face surveys tend to capture opinions from offline-inclined groups, whereas voluntary online surveys exclude those less active online.

Once the survey is designed, polling begins, and responses are aggregated by the polling organization. Respondents' answers are recorded and categorized, converting their opinions into structured data points, though these lack the nuances of human complexity. To protect privacy, data are often anonymized and aggregated, retaining key insights. After collection, raw data undergo cleaning and processing to minimize sampling biases and errors. Yet, any single poll can have random and systematic error sources, so combining multiple poll results helps mitigate these errors [@blumenthal2010polls].

Processed data are entered into structured datasets. Ideally, well-formatted polling data are stored as individual records, typically including fields like polling organization, sample size, survey dates, margin of error, candidate support percentages, and methodology description. On platforms like FiveThirtyEight, these fields are standardized for ease of comparison and analysis. Finally, FiveThirtyEight and similar platforms aggregate these structured records and present them to the public or allow non-commercial research use, with data usually updated as new polls are released. This enables analysts and the public to track changes in candidate support, analyze trends, and interpret results. An interesting phenomenon, however, is that as election day approaches, poll results often "converge" or "follow" the average, reducing polling error, with evidence suggesting some "final adjustments" in methodology may occur [@blumenthal2010polls].

@@

In researching the likelihood of Trump winning, we selected specific variables to focus on those most impactful for predicting approval rates. Here is an overview of data used for this analysis @tbl-view. Following are selected outcome variables and predictor variables, and explainations respectively.

### Outcome Variables
In this analysis of Donald Trump’s support rate in the 2024 U.S. presidential election, the outcome variable is Trump’s support rate (pct) because it is directly related to the election results. The primary research goal of this analysis is to predict Trump’s chances of winning in the 2024 election, so the support rate is the main metric we aim to model and forecast. The support rate here represents the proportion of respondents in a specific poll who support Trump. We seek to understand how the support rate is influenced by different factors, such as polling methodology, time, or geographical variables. By modeling the fluctuations in Trump’s support rate, we can infer his electoral performance, and assess the impact of various factors on his popularity. Understanding these dynamics provides insights into Trump’s standing relative to his opponents and supports our overall election prediction.

### Predictor Variables
In this analysis, the predictor variables are the factors that we believe can influence the support rate for Trump. These variables are chosen based on their potential to explain changes in support rate. Following are the predictor variables that will be included in our model, along with an explanation of their significance.

@@

**methodology**: Different survey methods (e.g., telephone interviews, online questionnaires) significantly impact the representativeness of the results, as they directly influence the characteristics of the sampled population. Some methods may introduce biases toward specific demographic groups.

@@

**days_since_Biden_Withdrawal**: This variable shows the difference between the start day of the poll and the number of days since Biden's withdrawal. Trump's main opponent, Kamala Harris, entered the race only after Biden withdrew. Major political events like Biden’s withdrawal can significantly impact voter approval rates, and tracking the time since this event helps assess both its short-term and long-term effects.

Other potential variables (such as voter demographics like gender, age, race, and income) could indeed influence voter approval rates, but we found that their correlation with presidential election outcomes was not very strong. Additionally, individual voter characteristics are not consistently available in most polls. Thus, we did not include these factors.

@@

### Relations Between Variables

@@

The plot on the left side in @fig-pct1 shows the average support rates for the top and bottom ten pollsters, highlighting each pollster’s most commonly used methodology. We can observe that the support rates between the top 5 and bottom 5 pollsters differ by about 10%, which helps us understand the bias introduced by certain pollsters’ influence on support rates, as some pollsters may tend to show higher or lower support rates. This chart also examines the different methodologies used by these pollsters, as varying survey methods (e.g., phone surveys or online panels) may affect the results by reaching different groups of voters. Focusing on these differences aids in assessing the reliability and bias of the data.

The plot on the left side in @fig-pct1 compares the average support rates across states with the national average, showing regional differences in voter preferences. Due to the Electoral College system in the U.S. presidential election, variations in state support rates can significantly impact the election outcome. This analysis highlights the differences between state and national polls, helping us understand Trump’s relative support in each state. Paying attention to geographic disparities improves the model's prediction accuracy.

@fig-pct2 displays the trend in support rates over time for the top five pollsters with the most observations. We chose these pollsters because a sufficient number of observations is needed to accurately capture trend changes and reduce the impact of extreme values. Each subplot represents a pollster and shows the change in Trump’s support rate since Biden’s withdrawal. This helps us observe the dynamic change in support rates following major events, and we can also see how different pollsters respond to the same event. Analyzing time trends aids in understanding shifts in public opinion, providing more time-sensitive support rate predictions.

```{r, message=FALSE, fig.width=6, fig.height=3}
#| label: fig-pct1
#| fig-cap: Support rates for Donald Trump which showed in different variables plot 1
#| echo: false

@@

p1 <- ggplot(selected_pollster_data, aes(x = reorder(paste(pollster, "(", count, ")", sep = ""), -avg_pct), y = avg_pct, fill = methodology)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ category, scales = "free_x") +
  labs(
    title = "Top 5 and Bottom 5 Avg Support Rate for Donald Trump",
    subtitle = "With Most Common Used Methodologies of Each Pollsters",
    x = "Pollster (Occurrence Count)",
    y = "Average Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 5, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 4, color = "#333333"),
    legend.text = element_text(size = 4, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
  ) +
  scale_fill_brewer(palette = "Set3") # assign colors to different methodologies

@@

p2 <- ggplot(selected_state_data, aes(x = reorder(paste(state, "(", count, ")", sep = ""), -avg_pct), y = avg_pct, fill = state)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ category, scales = "free_x") +
  labs(
    title = "Average Support Rate for Doanld Trump by states",
    subtitle = "National Average Support Rate Included",
    x = "State Names or National (Occurrence Count)",
    y = "Average Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 5, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
  ) +
  scale_fill_brewer(palette = "Set3") # assign colors to different states

p1 + p2
```

```{r, message=FALSE, fig.width= 5, fig.height= 3}
#| label: fig-pct2
#| fig-cap: Support rates for Donald Trump which showed in different variables plot 2
#| echo: false
### Identify the top 10 pollsters by occurrence
top_pollsters2 <- analysis_data %>%
  count(pollster, sort = TRUE) %>%
  top_n(5, n) %>%
  pull(pollster)

@@

# Step 2: Create the line plot with background distribution
ggplot(support_trend_5day, aes(x = grouped_days, y = avg_pct)) +
  geom_line(aes(color = pollster), linewidth = 0.7) + # Main line
  geom_jitter(data = data, aes(x = days_since_Biden_Withdrawal, y = pct), 
              color = "#a6cee3", alpha = 0.4, width = 0.5, height = 0.5) +
  scale_x_continuous(
    breaks = seq(min(analysis_data$days_since_Biden_Withdrawal), 
                 max(analysis_data$days_since_Biden_Withdrawal), 
                 by = 14)
  ) +
  labs(
    title = "Average Support Rate for Donald Trump Over Time",
    subtitle = "Days counted weekly since Biden's Withdrawal",
    x = "Days Since Biden's Withdrawal",
    y = "Average Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 4, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 4, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
  ) +
  facet_wrap(~ pollster, scales = "free_y")
```

# Model
Our model applies a hierarchical Bayesian approach, based on the framework proposed by Gelman et al. (2020).   This method builds a layered structure that combines polling data collected by different organizations using various methodologies with regional and temporal trends to more accurately predict support rates on Election Day.   Running within a Bayesian framework, our model employs a random walk prior to reflect the dynamically changing support rates following Biden's withdrawal in 2024.   This prior distribution is based on historical election trends and the latest polling data, allowing the model to continuously update as new data becomes available.

@@

## Model justification
### Model Prior Selection, Assumptions, and Validation
We based our prior choices on historical polling data and previous research in electoral studies, aiming to provide informative priors without overly constraining the model. For instance, we set a prior on the intercept as $\text{Normal} \sim (45, 5)$, assuming Trump’s support rate centers around 45%, with some uncertainty. Effect size priors (e.g., $\text{Normal} \sim (0, 5)$ for variables such as methodology and days_since_Biden_Withdrawal) reflect our expectation of moderate effects for these variables. This balance between informative priors and adaptability allows the model to accommodate data while providing a reasonable starting framework.

The model assumes that state-level differences and polling method effects are random and normally distributed. In reality, polling methods and state trends may exhibit non-normal behavior, especially during significant political events. Random effects for different polling agencies are introduced to account for biases associated with varying methodologies. However, we assume that each agency's methodology remains consistent over time, an assumption that may not hold if agencies adjust their sampling techniques. While the model effectively captures overall polling trends, it may not fully reflect subtle shifts within specific demographic subgroups (such as gender, race, socioeconomic status) since these details are not consistently available or may be obscured across poll results.

We implemented the model using the rstanarm package, leveraging MCMC sampling for Bayesian inference. To validate the model, we performed out-of-sample testing and calculated the RMSE (Root Mean Square Error), which resulted in 2.31. In the context of complex, socially influenced political surveys, this error level is considered acceptable. Convergence diagnostics show that R-hat values are close to 1, and the Effective Sample Size (ESS) is generally large (often over 400), indicating good independence among samples. Sensitivity analyses on priors indicate that slight adjustments do not significantly alter the predictions, demonstrating model stability.

We also considered a simpler linear regression model (excluding hierarchical structures) and a more complex model (incorporating additional economic data like state GDP and personal income). Through testing, we found that the simpler model could not adequately capture differences across polling agencies and states, while the more complex model added computational cost with minimal gains in accuracy. Consequently, we chose the current approach to balance flexibility, interpretability, and computational efficiency.

### Expectation of relationships between variables

We anticipate that different polling `methodology` may affect the estimation of support rates. While we cannot precisely determine which polling method might lean toward which demographic group, it is plausible to assume that methods like "online panels" or "phone surveys" might reach different segments of voters, thus introducing some bias into the support rate estimates. For instance, online surveys may engage more frequent internet users, while phone surveys may cover more traditional media users. This bias could lead to varying estimates of support for Trump.

For the variable `days_since_Biden_Withdrawal`, we hypothesize that, over time, Trump’s approval ratings may fluctuate due to shifts in the election landscape, policy discussions, and changes in voter sentiment. This hypothesis is based on the dynamic nature of polling data: as elections near their end, poll results typically converge, reflecting more accurate forecasts for election day (see Blumenthal's Polls, Forecasts, and Aggregators). We anticipate that the `days_since_Biden_Withdrawal` variable may exhibit a positive or negative trend, depending on the political climate and voter responses to Trump in particular periods.

Through the `is_national` variable, we distinguish between national and state-level polls, this variable enables the model to capture support rate differences arising from geographic and demographic differences. We expect that national polls, which typically have larger and more representative samples, are better suited to capturing overall trends, while state-level polls may reflect finer, localized biases and specific regional issues because national polls often encompass a broader and more balanced voter base, they provide a more stable forecast of overall voter trends. In contrast, due to geographic and demographic variations, state-level polls often exhibit greater fluctuations in error than national polls. These fluctuations stem from regional issues and diverse demographic characteristics, which may not align fully with national trends (Disentangling Bias and Variance in Election Polls).

Therefore, we included a random intercept `(1 | state)` for each state, with the intention that this random intercept would capture the differing impacts each state has on Trump’s support rate. Similarly, by incorporating `(1 + days_since_Biden_Withdrawal | pollster)`, we allow for variability in estimates by different pollsters over time. We anticipate that, due to differences in the stance or methodologies of these organizations, their estimates of Trump’s support rate in the period following Biden’s withdrawal will vary accordingly.

# Results
## Model Output
Our results are summarized in a table which is available [here](https://8360846bf92342d59862bf4ec3515643.app.posit.cloud/file_show?path=%2Fcloud%2Fproject%2Fpaper%2Fmodel_summary_table.html). This summary reveals the effects of different polling methods and organizations on our estimates. For instance, there is a small difference in estimates between data collected via online panels and phone surveys. The model adjusts for these differences by using a hierarchical structure, which accounts for inconsistencies caused by methodological and organizational biases. Specifically, variance parameters for different methods (such as online ads vs. phone surveys) indicate the different impacts these methods have on sampling coverage and voter behavior, helping the model balance these differences.

@@

modelsummary::modelsummary(
  list("First model" = first_model),
  statistic = "mad",
  fmt = 2,
  output = here::here("/cloud/project/paper/model_summary_table.html")
)
```

## Prediction Output
From @fig-results, the model predicts that nationwide support for Trump on Election Day, November 5, 2024, will center between 45% to 50% and shows moderate fluctuations. This variability is mainly attributed to national events and differences in polling methods. For instance, nationwide events such as economic shifts or unexpected political occurrences may influence voter sentiment, while differences in polling methods (such as phone surveys versus online panels) may introduce slight fluctuations in estimates due to differences in sample coverage.

At the state level, @fig-results shows predictions indicate significant variations between states. For example, support in Texas is considerably higher than in California, with a gap of over 10 percentage points, while swing states like Pennsylvania and Michigan display greater volatility. This result further validates the model’s robustness; the introduction of a random intercept for each state reflects systematic differences across states. This design aligns with data trends, suggesting that states with diverse geographic and demographic profiles exhibit unique political tendencies.

```{r, fig.width=5, fig.height=5}
#| echo: false
#| eval: true
#| label: fig-results
#| fig-cap: Prediction of average support for Donald Trump on 2024-11-05
#| warning: false
national_only <- prediction_data[prediction_data$state == "national", ]
all_states <- prediction_data[prediction_data$state != "national", ]

p4 <- ggplot(national_only, aes(x = state, y = predicted_support)) +
  geom_point(size = 3, color = "lightblue") +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, color = "lightblue") +
  labs(title = "Election Day National Trump Support Rate Prediction", y = "Predictied Support Rate (%)", x = "Area Type in US") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 6),
    plot.title = element_text(size = 8, face = "bold", color = "#333333"),
    axis.title = element_text(size = 7, color = "#333333"))

# Plot for state data
p5 <- ggplot(all_states, aes(x = state, y = predicted_support)) +
  geom_jitter(size = 3, color = "lightblue", width = 0.3) +  # Adds slight horizontal jitter to reduce overlap
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2, color = "lightblue", alpha = 0.6) +
  geom_smooth(aes(group = 1), method = "lm", color = "orange", se = FALSE) +  # Adds a trend line across all states
  geom_hline(yintercept = 50, color = "black", linetype = "dashed", size = 0.7) +  # Adds a reference line at 50%
  scale_y_continuous(breaks = seq(0, 70, by = 10)) +  # Formats y-axis as percentages with 1 decimal precision
  labs(title = "Election Day Trump Support Rate Prediction by State", y = "Predictied Support Rate (%)", x = "States") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5, angle = 90, hjust = 1),
    axis.text.y = element_text(size = 6),
    plot.title = element_text(size = 8, face = "bold", color = "#333333"),
    axis.title = element_text(size = 7, color = "#333333"))
